---
title: "Syllabus Day Slides"
author: "Kyle Bocinsky"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: "libs"
    chakra: "libs/remark-latest.min.js"
    css: ["default", "css/ath-slides.css", "css/ath-inferno-fonts.css", "css/animate.css"]
    seal: false
    anchor_sections: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.retina = 3, fig.align = "center")
```

```{r packages-data, include=FALSE}
library(tidyverse)
# library(ggdag)
```

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_xaringan_extra(c("tile_view","broadcast"))
xaringanExtra::use_share_again()
```

class: center middle main-title section-title-1

# FORS350 / GPHY488<br>(Forestry) Applications of GIS

.class-info[

.light[FORS350 / GPHY488<br>(Forestry) Applications of GIS<br>University of Montana<br>
WA Franke College of Forestry & Conservation
]

]

---


layout: false
name: class-details
class: center middle section-title section-title-6 animated fadeIn

# What I did during my winter break... as a map!

Draw a map (on paper, with your hands) that communicates something about what you did during winter break.

Think about how your map tells a story... what would one of your colleagues or I glean from your map? How did you use text, color, and other map elements to convey your story?

**Please write your name on top of your map.**

---


layout: false
name: class-details
class: center middle section-title section-title-6 animated fadeIn

# What I did during my winter break... as a map!

Trade maps with the person next to you and try to interpret what your partner's map is trying to convey.

What elements are confusing, or clear, about your partner's map? How would you tell the same story differently?

**Discuss.**

---

#  Course Website

.box-8.medium[![](https://quickchart.io/qr?text=https%3A%2F%2Fumtgis.github.io%2Fapps%2F&size=300)]

.box-inv-6.sp-after[https://umtgis.github.io/apps/]

---

#  Final Notes

.box-inv-2.medium[Turn in your Maps! (10 pts)]

--

.box-inv-4.medium[Respond to my email! (10 pts)]

--

.box-inv-6.medium[Course Waitlist]

--

.box-inv-8.medium[Lab Today]

.box-8.sp-after-half[(if you want assistance setting up a personal computer)]
