---
title: "How is GIS applied to the scientific method?"
date: "2024-01-23"
---

```{r}
#| label: slides-videos
#| echo: false
#| include: false
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- ""

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "Introduction", "1", TRUE,
  "", "Motivation", "motivation", FALSE,
  "", "Review", "review", FALSE,
  "", "Key concepts", "keyconceptsexperimental", FALSE
)
```

## Preparation
Please read [Chapter 1: Maps, Society, and Technology](https://open.lib.umn.edu/mapping/chapter/1-maps-society-and-technology/) from [Mapping, Society, and Technology](https://open.lib.umn.edu/mapping/) [@Manson:2017].

## Content

- The Scientific Method
- The Geographic Approach
- Geographic Representation

## Slides

The slides for today are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02-slides")
# slide_tabs(slide_details, "/slides/01-slides.html")
```

```{r show-slides-xaringan, echo=FALSE}
xaringanExtra::embed_xaringan(
  url = "/slides/02-slides.html",
  ratio = "16:9"
)
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of slide-specific commands.
:::
