---
title: "Whence Geospatial Data I: All data are spatial"
date: "2024-02-06"
---

```{r}
#| label: slides-videos
#| echo: false
#| include: false
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- ""

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "Introduction", "1", TRUE,
  "", "Motivation", "motivation", FALSE,
  "", "Review", "review", FALSE,
  "", "Key concepts", "keyconceptsexperimental", FALSE
)
```


## Preparation
Please read [Chapter 8. Surveillance](https://open.lib.umn.edu/mapping/chapter/8-surveillance/){target="_blank"} from [Mapping, Society, and Technology](https://open.lib.umn.edu/mapping/){target="_blank"} [@Manson:2017].

## Content

- Geospatial Authorities: Who creates, distributes, and maintains authoritative geodata?
  - [FAIR Guiding Principles for scientific data management and stewardship](https://www.go-fair.org/fair-principles/){target="_blank"} 
- Spatial metadata and data provenance
- Some geodata resources:
  - [The US Geospatial Platform](https://www.geoplatform.gov){target="_blank"}
  - [USGS National Geospatial Program](https://www.usgs.gov/programs/national-geospatial-program){target="_blank"}
    - [3D National Topography Model (3DNTM)](https://www.usgs.gov/programs/national-geospatial-program/3d-national-topography-model){target="_blank"}
      - [3D Elevation Program (3DEP)](https://www.usgs.gov/3d-elevation-program){target="_blank"}
      - [3D Hydrography Program (3DHP)](https://www.usgs.gov/3d-hydrography-program){target="_blank"}
    - [U.S. Board on Geographic Names](https://www.usgs.gov/us-board-on-geographic-names){target="_blank"}
      - [Geographic Names Information System (GNIS)](https://edits.nationalmap.gov/apps/gaz-domestic/public/search/names){target="_blank"}
    - [The National Map](https://www.usgs.gov/programs/national-geospatial-program/national-map){target="_blank"}
    - [TNM Supporting Themes](https://www.usgs.gov/programs/national-geospatial-program/supporting-themes){target="_blank"}
  - [Multi-Resolution Land Characteristics (MRLC) Consortium](https://www.mrlc.gov){target="_blank"}
  - [US Census Tiger Database](https://www.census.gov/programs-surveys/geography/guidance/tiger-data-products-guide.html){target="_blank"}
  - [ESRI Living Atlas of the World](https://livingatlas.arcgis.com/en/home/){target="_blank"}
  - [Montana Spatial Data Infrastructure (MSDI)](https://msl.mt.gov/geoinfo/data/msdi/){target="_blank"}
  - [Missoula Maps](https://missoulamaps-cityofmissoula.hub.arcgis.com){target="_blank"}
- Getting data into ArcGIS Pro

## Slides

The slides for today are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/20240206-whence_geospatial_data_I")
# slide_tabs(slide_details, "/slides/01-slides.html")
```
