---
title: "Introduction"
date: "2024-01-18"
---

```{r}
#| label: slides-videos
#| echo: false
#| include: false
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- ""

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "Introduction", "1", TRUE,
  "", "Motivation", "motivation", FALSE,
  "", "Review", "review", FALSE,
  "", "Key concepts", "keyconceptsexperimental", FALSE
)
```

## Content

- What I did during my winter break... as a map!
- Introductions
- [Syllabus](/syllabus), [schedule](/schedule/), and learning objectives

## Winter Break Map (20 minutes)
Draw a map (on paper, with your hands) that communicates something about what you did during winter break. Think about how your map tells a story... what would one of your colleagues or I glean from your map? How did you use text, color, and other map elements to convey your story? **Please write your name on top of your map.**

After you complete your map in 20 minutes, trade maps with the person next to you and try to interpret what your partner's map is trying to convey. What elements are confusing, or clear, about your partner's map? How would you tell the same story differently? Discuss.

## Syllabus Intro and Discussion
Familiarize yourself with the [syllabus](/syllabus), [schedule](/schedule/), [content](/content/), and [evaluations](/evaluations/) pages for this class.

## Slides

The slides for today are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/01-slides")
# slide_tabs(slide_details, "/slides/01-slides.html")
```

```{r show-slides-xaringan, echo=FALSE}
xaringanExtra::embed_xaringan(
  url = "/slides/01-slides.html",
  ratio = "16:9"
)
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of slide-specific commands.
:::
