---
title: "Quantitative maps: Choropleth Mapping"
date: "2024-03-05"
---

```{r}
#| label: slides-videos
#| echo: false
#| include: false
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- ""

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "Introduction", "1", TRUE,
  "", "Motivation", "motivation", FALSE,
  "", "Review", "review", FALSE,
  "", "Key concepts", "keyconceptsexperimental", FALSE
)
```

## Preparation
Please read [Chapter 5. Simplification](https://open.lib.umn.edu/mapping/chapter/5-simplification/){target="_blank"} from [Mapping, Society, and Technology](https://open.lib.umn.edu/mapping/){target="_blank"} [@Manson:2017].


## Content

Most of the content from todays lecture comes from the incredible paper *The misuse of colour in science communication* by Fabio Crameri, Grace E. Shephard & Philip J. Heron:

> Crameri, F., Shephard, G.E. & Heron, P.J. The misuse of colour in science communication. *Nature Communications* 11, 5444 (2020). <https://doi.org/10.1038/s41467-020-19160-7>

## Slides

The slides for today are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/20240305_color_and_choropleths")
# slide_tabs(slide_details, "/slides/01-slides.html")
```

```{r show-slides-xaringan, echo=FALSE}
xaringanExtra::embed_xaringan(
  url = "/slides/20240305_color_and_choropleths.html",
  ratio = "16:9"
)
```

<!-- :::{.callout-tip} -->
<!-- **Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of slide-specific commands. -->
<!-- ::: -->
