---
title: "Final Group Projects: Overview, Topics, and Q&A"
date: "2024-02-29"
---

```{r}
#| label: slides-videos
#| echo: false
#| include: false
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- ""

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "", "Introduction", "1", TRUE,
  "", "Motivation", "motivation", FALSE,
  "", "Review", "review", FALSE,
  "", "Key concepts", "keyconceptsexperimental", FALSE
)
```

## Preparation
We will be discussing the Final Group Projects today in class. The assignment can be found here: [Group GIS Project](/evaluations/final-project.html)


<!-- ## Content -->

<!-- - Spheroids, Ellipsoids, and Geoids (Oh My!) -->
<!-- - Projecting Earth -->
<!-- - Scale -->
<!-- - Geodatabases (and friends!) -->

## Slides

The slides for today are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/20240229-final-project")
# slide_tabs(slide_details, "/slides/01-slides.html")
```

```{r show-slides-xaringan, echo=FALSE}
xaringanExtra::embed_xaringan(
  url = "/slides/20240229-final-project",
  ratio = "16:9"
)
```

<!-- :::{.callout-tip} -->
<!-- **Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of slide-specific commands. -->
<!-- ::: -->
